---
title: "Assignment 4"
author: "Nathan Angell"
date: "9/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## problem 1
#### Install tidyverse package and dyplr
```{r}
library(tidyverse)
library(dplyr)
```

## problem 2
#### Use read_csv function to import the US Covid 19 data at link. Don’t forget to import tidyverse (library(tidyverse)) so that you can use read_csv.
```{r}
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')
```

## problem 3
#### Use the following codes to change the type of the date column to date
```{r}
library(lubridate)
df$date <- ymd(df$date)
```
## problem 3a
#### Use the below codes to create month, weekday and monthday variables
```{r}
df$month = month(df$date)

# day of the week
df$weekday = wday(df$date)

# day of the month
df$monthday <- mday(df$date)
```

## problem 4
#### The function case_when is a good option to create a new variable from existing variable.
```{r}
# create a new variable, daily_death, from deathIncrease variable deathIncrease is the number of daily new death by Covid19
df$daily_death <- case_when(
  df$deathIncrease <3 ~ 'low',
  df$deathIncrease <=14 ~ 'medium',
  TRUE ~ 'high'
) #if deathIncrease less than 3), medium (deathIncrease from 3 to 14), and high (deathIncrease more than 14)
```
## problem 4a
#### Create variable month2 that takes three values: early_month (day of the month from 1-10), mid_month (day of the month from 11-20), and end_month (day of the month > 20)
```{r}
df$month2 <- case_when(
  df$month <=10 ~ 'early_month',
  df$month >20 ~ 'mid_month',
  TRUE ~ 'end_month'
)
```
## problem 4b
#### Create variable weekend that takes two values: 1 if it’s Saturday or Sunday or 0 otherwise.
```{r}
df$weekend <- case_when(
  df$weekday >= 6 ~ 1,
  TRUE ~ 0
)
```
## problem 5
#### Use the select function to deselect the column totalTestsViral from the data.
```{r}
df1 <- select(df, -totalTestsViral)
```
##problem 6
#### Pipe operator offers another way to write R codes. Many times, it makes the codes more readable. Pipe works very well with all the tidyverse packages. Refer to these slides (slide 15, 16, 17 and 18) to rewrite the below codes using pipe operator
```{r}
# no pipe
x <- c(1:10)

# take squre root
x %>% sqrt

# sum it
x %>% sqrt %>% sum()

#log 
x %>% sqrt %>% sum() %>% log()

#log base 2 of 16
16 %>% log(2)
```
## problem 7
#### This combo is used when you want to apply a function/calculation to different groups of the data. For example, to calculate the average number of cases (positiveIncrease) by dataQualityGrade, we use: 
```{r}
df %>% 
  group_by(weekday) %>% 
  summarise(mean(positiveIncrease))
```
## problem 7a
#### Calculate the median number of cases (positiveIncrease) by month
```{r}
# we have to summarise because we need to combine the data together
df %>% 
  group_by(month) %>% summarise((median(positiveIncrease)))
```
## problem 7b
#### Calculate the average number of cases (positiveIncrease) by month2
```{r}
df %>% 
  group_by(month2) %>% summarise((mean(positiveIncrease)))
```
## problem 7c
#### Calculate the median number of cases (positiveIncrease) by weekend
```{r}
df %>% 
  group_by(weekend) %>% summarise((median(positiveIncrease)))
```
## problem 8
#### to calculate the average number of cases (positiveIncrease) in January and February separately, we use:
```{r}
df %>% 
  filter(month==1|month==2) %>% 
  group_by(month) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```
## problem 8a
#### Calculate the median number of cases (positiveIncrease) on the weekend by month in October and November 2020.
```{r}
df %>% 
  filter(month==10|month==11) %>% 
  group_by(month) %>% 
  summarise(positve_increase = median(positiveIncrease))
```
## problem 8b
#### Calculate the average number of death at different periods of a month (month2 variable) in Fall 2020
```{r}
df %>% 
  filter(month2=='early_month'|month2=='end_month') %>% 
  group_by(month2) %>% 
  summarise(positve_increase = mean(positiveIncrease))
```
## problem 8c
#### Compare the average number of hospitalizations between weekdays and weekends in Summer 2020
```{r}
df %>% 
  filter(weekend==1|weekend==0) %>% 
  group_by(weekend) %>% 
   mean(hospitalized)
```
## problem 8d
#### Find the median cases by weekdays in Rhode Island in 2021
```{r}
# need to do this stuff before 14
df_assignment3 <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv') 
df_assignment3$date <- as.Date(df_assignment3$date)
df_assignment3$Weekdays <- weekdays(df_assignment3$date)
df_assignment3_ <- df_assignment3[(df_assignment3$state == 'Rhode Island'),]

#actual assignment
df_assignment3_1 <- group_by(df_assignment3_, Weekdays)
summarise(df_assignment3_1, median_cases = median(cases))
```
## problem 8e
#### Compare the median cases in Rhode Island in June, July, August and September in 2021.
```{r}
df_assignment3_2 <- group_by(df_assignment3_, date>='2021-06-01')
summarise(df_assignment3_2, median_cases = median(cases))
```
## problem 9
#### Use the arrange function to find a month that has the highest number of deaths on the weekend.
```{r}
df %>% filter(weekend == 1) %>% summarise(average_deaths = mean(death)) %>% arrange(average_deaths)

```