---
title: "assignment 3"
author: "Nathan Angell"
date: "9/15/2021"
output: html_document
---
## problem 1

```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28,34,29,42),
                Name = c("Tom", "Jack", "Steve", "Ricky"))
```
## problem 2
### use the read.csv to get the df into your system
```{r}
df = read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
```

## problem 3
### How many columns and rows the data have?
```{r}
str(df)
# there are 30,869 rows and 5 columns from what I saw from this code
```
## problem 4
### How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?
```{r}
colSums(is.na(df))
# there are no missing values
```
## problem 5
### What is the class of the date column? Change the date columns to date type using the as.Date function. Show the new class of the date column
```{r}
class(df$date)

# you still need to knit the assignment as html

```




