---
title: "Assignment 11"
author: "Nathan Angell"
date: "10/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## install packages and import data
- Set seed to be 2020.
- The target variable is diabetes
- Partition the data into 80% training and 20% testing.
```{r}
library(mlbench)
library(caret)
data(PimaIndiansDiabetes)
df <- PimaIndiansDiabetes

#seed
set.seed(2020)

#partition data, diabetes is the target variable
splitIndex <- createDataPartition(df$diabetes, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

## 2. Tuning Decision Tree: Use cross-validation with 10 k-folds to find the maxdepth with the greatest accuracy. Plot the accuracy associated with different maxdepths against the maxdepths. The range to search for maxdepth is from 1 to 10.

```{r}
# Decide the range of the maxdepth to search for the best
tuneGrid = expand.grid(maxdepth = 1:10)
# Tell caret to do Approach 2, i.e. Cross-Validation
trControl = trainControl(method = "cv",
                         number = 10)
# Do Approach 2 
tree_approach2 <- train(diabetes~., data=df_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
```

## 3. Make the final decision to select the maxdepth for your decision tree. Is your selected maxdepth the same as the maxdepth found in 2?
According to the plot, my decision is to have a depth of 3. It is not the best model,accuracy wise, but I think it is a model that is usable for other different types of situations If you make it only pertain to this specific dataset, it is not going to be a model that can go to other datasets, it will be too particular. 
```{r}
plot(tree_approach2)
```
## 4. Calculate the accuracy of your decision tree (the decision tree with your selected maxdepth in 3) on the test data.
```{r}
decision_tree <- train(diabetes~., data=df_train, 
                method = "rpart2",
                maxdepth=3)
pred <- predict(decision_tree, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$diabetes, positive = "pos")
cm$overall[1]
```






