---
title: "Assignment 8"
author: "Nathan Angell"
date: "10/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Making a bar race chart

#### install packages and download dataset
1. Install two packages gganimate and gifski then restart Rstudio. Use the WHO’s dataset. Make a top-10 bar race by months between countries on the number of deaths by Covid 19.
```{r}
library(gganimate)
library(gifski)
library(tidyverse)
library(dplyr)
library(lubridate)

#df
df <- read_csv("WHO-COVID-19-global-data.csv")


#bar race chart
df$month <- month(df$Date_reported)
d1 <- df %>% group_by(month, Country) %>% summarise(mean = mean(Cumulative_deaths))
d2 <- d1 %>% group_by(month) %>% mutate(rank=rank(-mean)) %>% ungroup() #-mean --> largest to small
d3 <- d2 %>% filter(rank <= 10)
a1 <- d3 %>% ggplot(aes(x=rank, y=mean, group=Country, fill=Country, label=Country)) + geom_col()+
    geom_text(aes(y = mean, label = Country), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Month {closest_state}', x='', y='Total Number of Deaths', fill='Country')+
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(month)+
    ease_aes("cubic-in-out")
animate(a1, nframes = 400)
```

#### 2. Make another bar race chart using that dataset.
```{r}
d4 <- df %>% group_by(month, Country) %>% summarise(mean = mean(Cumulative_cases))
d5 <- d4 %>% group_by(month) %>% mutate(rank=rank(-mean)) %>% ungroup() #-mean --> largest to small
d6 <- d5 %>% filter(rank <= 10)
a1 <- d6 %>% ggplot(aes(x=rank, y=mean, group=Country, fill=Country, label=Country)) + geom_col()+
    geom_text(aes(y = mean, label = Country), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Month {closest_state}', x='', y='Total Number of Cases', fill='Country')+
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(month)+
    ease_aes("cubic-in-out")
animate(a1, nframes = 400)
```

#### 3. Make a bar race using a dataset of your own interest. You may use the dataset that we use in class (in code below) but you should make a different bar race from ones in the slides.
```{r}
# import df1
df1 <- read_csv('all-states-history.csv')

df1$month <- month(df1$date)
d7 <- df1 %>% group_by(month, state) %>% summarise(mean = mean(totalTestResults))
d8 <- d7 %>% group_by(month) %>% mutate(rank=rank(-mean)) %>% ungroup() #-mean --> largest to small
d9 <- d8 %>% filter(rank <= 10)
a1 <- d9 %>% ggplot(aes(x=rank, y=mean, group=state, fill=state, label=state)) + geom_col()+
    geom_text(aes(y = mean, label = state), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Month {closest_state}', x='', y='Total Number of Test Results', fill='state')+
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(month)+
    ease_aes("cubic-in-out")
animate(a1, nframes = 400)

```





