---
title: "Assignment 14"
author: "Nathan Angell"
date: "11/22/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Pick a keyword or hashtag. Download the data associated with the keyword/hashtag. Plot at least 10 plots to visualize the data associated with the keyword/hashtag. All plots should have titles and captions.

```{r}
#library(rtweet) 
library(tidytext)
library(ggpubr) 
library(tidyverse) 
library(knitr)
library(lubridate)

keyword_search = '#microsoft'

#df <- search_tweets(q = keyword_search, 
                        #n = 18000,
                        #include_rts = FALSE,
                        #`-filter` = "replies",
                        #lang = "en") %>% 
  #mutate(created_at = ymd_hms(format(created_at, tz = "US/Eastern")))

#write_csv(df, 'twitter_data_ole.csv')

```

```{r}
df <- read.csv("twitter_data_ole.csv")
```


```{r}
df %>% 
  unnest_tokens(output = hashtag, input = text, token = "tweets") %>%
  filter(!hashtag %in% c("#microsoft"), str_detect(hashtag, "^#")) %>% 
  count(hashtag, sort = TRUE) %>%
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(hashtag, n)))+
  geom_col()+
  labs(x = 'Frequency', y = '')+
    labs(x='Count', 
       y = 'Hashtags', 
       title = 'Count vs Hashtags', 
       caption = 'Top 10 hashtags.')
```

```{r}
library(RColorBrewer)

pal <- brewer.pal(8,"Dark2")
library(wordcloud) 

df %>% 
  unnest_tokens(output = hashtag, input = text, token = "tweets") %>%
  filter(!hashtag %in% c("#microsoft"), str_detect(hashtag, "^#")) %>%
  count(hashtag, sort = TRUE) %>%
  mutate(hashtag = str_remove(hashtag, '#')) %>% 
  filter(hashtag != 'mufc') %>% 
  with(wordcloud(hashtag, n, random.order = FALSE, max.words = 50, colors = pal))
```

```{r}
#top five favorites
df %>% 
  arrange(-favorite_count) %>%
  head(5) %>% 
  select(favorite_count, text, favorite_count)
```

```{r}
#screen name
df %>% 
  count(screen_name, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(screen_name, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'Screen name', 
       title = 'Count vs Screen Name', 
       caption = 'Top 5 Screen Name')
```

```{r}
df %>% 
  unnest_tokens(mentions, text, "tweets", to_lower = FALSE) %>%
  filter(str_detect(mentions, "^@")) %>%  
  count(mentions, sort = TRUE) %>%
  head(10)

df %>% 
  count(source, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(source, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'Source', 
       title = 'Count vs Source', 
       caption = 'Top 5 Sources')
```

```{r}
df %>% 
  filter(!is.na(country)) %>% 
  count(country, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(country, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'County', 
       title = 'Count vs Country', 
       caption = 'Top 5 Country')
```

```{r}
df %>% 
  filter(!is.na(location), !location=='') %>% 
  count(location, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(location, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'Location', 
       title = 'Count vs Location', 
       caption = 'Top 5 Location')
```

```{r}
df %>% 
  select(screen_name, favourites_count) %>% 
  filter(!is.na(favourites_count)) %>% 
  group_by(screen_name) %>% 
  summarise(average_fav = mean(favourites_count)) %>% 
  arrange(-average_fav) %>% 
  head(5) %>% 
  ggplot(aes(x=average_fav, y = reorder(screen_name, average_fav)))+
  geom_col()+
    labs(x='Count', 
       y = 'User', 
       title = 'Count vs User', 
       caption = 'Top 5 Users')
```

```{r}
df %>%
    unnest_tokens(input = text, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(word, sort = TRUE) %>%
    inner_join(get_sentiments("nrc")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(y= reorder(sentiment, n), x = n))+geom_col()+
    labs(y='Relative Frequency', x ='')+
    labs(x='Count', 
       y = 'Sentiment Analysis', 
       title = 'Count vs Sentiment Analysis', 
       caption = 'Top 10 Sentiment Analyses')
```
```{r}
df %>%
    unnest_tokens(input = text, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(word, sort = TRUE) %>%
    inner_join(get_sentiments("bing")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(y= reorder(sentiment, n), x = n))+geom_col()+
    labs(x='Count', 
       y = 'Sentiment Analysis', 
       title = 'Count vs Sentiment Analysis', 
       caption = 'The general Sentiment Analysis')
```
#### Choose a location then pick a trending keyword/hashtag in the location. Download the data associated with the keyword/hashtag. Plot at least 10 plots to visualize the data associated with the keyword/hashtag. All plots should have titles and captions.
```{r}
#trends_available()
```
```{r}
#df <- get_trends('Quebec')
```

```{r}
## lookup users by screen_name or user_id
users <- c("KimKardashian", "justinbieber", "taylorswift13",
           "espn", "JoelEmbiid", "cstonehoops", "KUHoops",
           "upshotnyt", "fivethirtyeight", "hadleywickham",
           "cnn", "foxnews", "msnbc", "maddow", "seanhannity",
           "potus", "epa", "hillaryclinton", "realdonaldtrump",
           "natesilver538", "ezraklein", "annecoulter")

#famous_tweeters <- lookup_users(users)

## preview users data
#famous_tweeters
```

```{r}
#tweets_data(famous_tweeters)
```
```{r}
keyword_search = 'Lebron'

#df <- search_tweets(q = keyword_search, 
                       # n = 18000,
                       # include_rts = FALSE,
                       # `-filter` = "replies",
                       # lang = "en") %>% 
 # mutate(created_at = ymd_hms(format(created_at, tz = "US/Eastern")))

#write_csv(df, 'twitter_data_lebron.csv')
```

```{r}
#we noticed the that lebron is one of the top people based upon the above code 
df <- read_csv("twitter_data_lebron.csv")

df %>% 
  unnest_tokens(output = hashtag, input = text, token = "tweets") %>%
  filter(!hashtag %in% c("#lebron","#lebronjames"), str_detect(hashtag, "^#")) %>%
  count(hashtag, sort = TRUE) %>%
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(hashtag, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'Hashtags', 
       title = 'Count vs Hashtags', 
       caption = 'Top 10 Hashtags')
```
```{r}
library(RColorBrewer)

pal <- brewer.pal(8,"Dark2")
library(wordcloud) 

df %>% 
  unnest_tokens(output = hashtag, input = text, token = "tweets") %>%
  filter(!hashtag %in% c("#lebron","#lebronjames"), str_detect(hashtag, "^#")) %>%
  count(hashtag, sort = TRUE) %>%
  mutate(hashtag = str_remove(hashtag, '#')) %>% 
  filter(hashtag != 'mufc') %>% 
  with(wordcloud(hashtag, n, random.order = FALSE, max.words = 50, colors = pal))
```

```{r}
df %>% 
  count(screen_name, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(screen_name, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'Screen name', 
       title = 'Count vs Screen Name', 
       caption = 'The top five screen name')
```

```{r}
df %>% 
  count(source, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(source, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'Sources', 
       title = 'Count vs Sources', 
       caption = 'The general Sentiment Analysis')
```

```{r}
df %>% 
  filter(!is.na(country)) %>% 
  count(country, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(country, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'Sources', 
       title = 'Count vs Sources', 
       caption = 'The general Sentiment Analysis')
```

```{r}
df %>% 
  filter(!is.na(location), !location=='') %>% 
  count(location, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(location, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'County', 
       title = 'Count vs County', 
       caption = 'The count for each of the top counties')
```

```{r}
df %>% 
  select(screen_name, favourites_count) %>% 
  filter(!is.na(favourites_count)) %>% 
  group_by(screen_name) %>% 
  summarise(average_fav = mean(favourites_count)) %>% 
  arrange(-average_fav) %>% 
  head(5) %>% 
  ggplot(aes(x=average_fav, y = reorder(screen_name, average_fav)))+
  geom_col()+
    labs(x='Count', 
       y = 'Accounts', 
       title = 'Count vs Accounts', 
       caption = 'Top Accounts with the highest favorites')
```

```{r}
df %>% 
  unnest_tokens(output = word, input = text, token = "words") %>%
  anti_join(get_stopwords()) %>% 
  filter(!word %in% c('https', 't.co')) %>% 
  count(word, sort = TRUE) %>%
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(word, n)))+
  geom_col()+
    labs(x='Count', 
       y = 'Tweet', 
       title = 'Count vs Tweet', 
       caption = 'Highest text in a specific tweet')
```

```{r}
df %>%
    unnest_tokens(input = text, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(word, sort = TRUE) %>%
    inner_join(get_sentiments("nrc")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(y= reorder(sentiment, n), x = n))+geom_col()+
    labs(x='Count', 
       y = 'Sentiment', 
       title = 'Count vs Sentiment', 
       caption = 'Sentiment analysis for tweets')
```

```{r}
df %>%
    unnest_tokens(input = text, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(word, sort = TRUE) %>%
    inner_join(get_sentiments("bing")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(y= reorder(sentiment, n), x = n))+geom_col()+
    labs(x='Count', 
       y = 'Sentiment Analysis', 
       title = 'Count vs Sentiment Analysis', 
       caption = 'Average sentiment analysis for all tweets')
```

