---
title: "Assignment 15"
author: "Nathan Angell"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidytext)
library(ggpubr) 
library(tidyverse) 
library(knitr)
library(lubridate)
```


### 1. Write a function to calculate the age of a person. Give an example to test your function.
- Input: a number, x, (year born)
- Output: Print out “You are r age”. Where r is the age of the person, i.e. 2020 - x.
```{r}
age <- function(x)
{
  r = 2021-x
  print(paste0('You are ', r, ' years old.'))
}

age(1999)
```


### 2. Write the following function. Give an example to test your function.
- Input: a number
- Output: print out: “Your input is an even number!” if the number is not even, or “You input an odd number!” otherwise.
```{r}
num <- function(x)
{
  if((x %% 2) == 0) 
  {
  print(paste(x,"is Even"))
  } else 
  {
  print(paste(x,"is Odd"))
  }
}

num(3)

num(4)
```

### 3. Write the following function. Give an example to test your function.
- Input: a numeric vector
- Output:
  - if the input vector has missing values: return the input vector with missing values replaced by mean
  - if the input vector has no missing value: return the same input vector
```{r}
mean_impute <- function(x)
{
  if(sum(is.na(x))>0)
  {
    # Find the mean of x
    mean_of_x <- mean(x, na.rm = TRUE)
    # Replace the missing with mean
    library(tidyr)
    x <- replace_na(x, mean_of_x) 
  }
return(x)    
}

x <- c(1,2,5,3,NaN,6,7,NaN)

mean_impute(x)
```
  
### 4. Write the following function. Give an example to test your function.
- Input: a vector x
- Output: The vector x where the missing values replaced by the mean (if x is numeric) or the mode (if x is non-numeric). If x does not have missing value, return the same vector x.

```{r}
impute <- function(x)
{
  if(is.numeric(x))
  {
    # Find the mean of x
    mean_of_x <- mean(x, na.rm = TRUE)
    # Replace the missing with mean
    x <- replace_na(x, mean_of_x) 
  }
  
  else {
    library(tidyr)
     # Find the mean of x
     mode_of_x <- names(sort(-table(x)))[1]
    # Replace the missing with mean
    x <- replace_na(x, mode_of_x) 
    
  }
return(x)    
}

x <- c(1,NA,3,NA,5,8)
impute(x)

x <- c('a','b','a','a','z',NA,'d', NA)
impute(x)
```

### 5. Write the following function. Give examples to test your function.
- Input: A data frame of two variables x and y
- Output:
  - A boxplot of x by y if x is numeric and y is non-numeric
  - A boxplot of y by x if y is numeric and x is non-numeric
  - print out ‘This function cannot visualize your data’ otherwise
  
```{r}
viz <- function(d)
{
  if(is.numeric(d[[1]])& is.character(d[[2]]))
  {
    d %>% ggplot(aes(x = d[[1]], y = d[[2]]))+
      geom_point()+
      labs(x = names(d)[1], y = names(d)[2])
  }
  else if ((is.numeric(d[[2]])& is.character(d[[1]])))
  {
    d %>% ggplot(aes(x = d[[2]], fill = d[[1]]))+
      geom_bar(position = 'dodge')+
      labs(x = names(d)[2], fill = names(d)[1])
  }
  else 
  {
    print('This function cannot visualize your data.')
  }
}

df <- read_csv("titanic.csv")

d <- df %>% select(Age, Embarked)
viz(d)

d <- df %>% select(Embarked, Age)
viz(d)
```
  
### 6. Combine the function in Question 5 and the function in this example (Link) to have a function that can plot all possible cases (2 categorical, 2 numeric, 1 numeric and 1 categorical).

```{r}
viz <- function(d)
{
  if(is.numeric(d[[1]])& is.character(d[[2]]))
  {
    d %>% ggplot(aes(x = d[[1]], y = d[[2]]))+
      geom_point()+
      labs(x = names(d)[1], y = names(d)[2])
  }
  else if ((is.numeric(d[[2]])& is.character(d[[1]])))
  {
    d %>% ggplot(aes(x = d[[2]], fill = d[[1]]))+
      geom_bar(position = 'dodge')+
      labs(x = names(d)[2], fill = names(d)[1])
  }
  
  else if(is.numeric(d[[1]])&is.numeric(d[[2]]))
  {
    d %>% ggplot(aes(x = d[[1]], y = d[[2]]))+
      geom_point()+
      labs(x = names(d)[1], y = names(d)[2])
  }
  else if (!(is.numeric(d[[1]])|is.numeric(d[[2]])))
  {
    d %>% ggplot(aes(x = d[[1]], fill = d[[2]]))+
      geom_bar(position = 'dodge')+
      labs(x = names(d)[1], fill = names(d)[2])
  }
  
  else 
  {
    print('This function cannot visualize your data.')
  }
}

d <- df %>% select(Age, Fare)
viz(d)

df$Pclass <- factor(df$Pclass)
d <- df %>% select(Sex, Pclass)
viz(d)
```


### 9. Write the following function. Give examples to test your function.
- Input:
  - a data frame that has a text column
  - the name of the text column in the data
  
- Output: the word cloud plot of the text column
```{r}
word_frequency <- function(text_df, text_col) {
  library(wordcloud)
  text_df %>% 
  unnest_tokens(input = text_col, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(word, sort = TRUE) %>% 
  head(5) %>% 
  ggplot(aes(x = n, y = reorder(word, n))) +
  geom_col() + theme(axis.text.y = element_text(size = 40))+ 
  labs(y = '', x = 'Frequency')
}

df <- read_csv('twitter_data_lebron.csv')
word_frequency(text_df = df, text_col = 'text')
```





