---
title: "Assignment 6"
author: "Nathan Angell"
date: "10/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Install tidyverse package and dyplr
```{r}
library(tidyverse)
library(dplyr)
```

#### import data
```{r}
df <- read_csv("titanic.csv")
```

## 1
#### For one continuous variable: Density, Histogram and Boxplot
```{r}
#density
#this basically says most of the data is in between 10-40
df %>% ggplot()+
  geom_density(mapping = aes(x = Age))

#histogram
#shows that fares are between 0-100 dollars
df %>% ggplot()+
  geom_histogram(mapping = aes(x = Fare))

#boxplot
#shows the median, which is around 20 or so dollars. with a couple of outliers. 
df %>% ggplot()+
  geom_boxplot(mapping = aes(x = Fare))
```
## 2
#### create one cateorical barplot
```{r}
#most people are in the 'S' embarked
df %>% ggplot()+
  geom_bar(mapping = aes(x = Embarked))
```

## 3
#### For one two continuous variables: Scatter Plot, Line Plot and Smooth Plot
```{r}
# scatter plot
# most people were below 50 for all ages. with maybe 100 or so above that shreshold for the fare
df %>% ggplot()+
  geom_point(mapping = aes(x = Age, y= Fare))

# line plot
# most people in the pclass of 3, have more siblings
df %>% ggplot()+
  geom_line(mapping = aes(x = Pclass, y= SibSp))

# smooth plot
# most people in the pclass of 3, have more siblings
df %>% ggplot()+
  geom_smooth(mapping = aes(x = Pclass, y= SibSp))
```

## 4
#### For one continuous + one categorical variables: Density and boxPlot
```{r}
#density
#the people that have NA's have higher ages
df %>% ggplot()+
  geom_density(mapping = aes(x = Age, color= Embarked))

# boxplot
#the people that have NA's have higher ages
df %>% ggplot()+
  geom_boxplot(mapping = aes(x = Age, color= Embarked))
```

## 5
#### For two categorical variables: barplot
```{r}
df %>% ggplot()+
  geom_bar(mapping = aes(x = Embarked))

df %>% ggplot()+
  geom_bar(mapping = aes(x = Parch))

```

# Questions on Covid Vaccination Data

## import the correct libaries
```{r}
library(tidyverse)
df <- read_csv('https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/vaccine_data_us_timeline.csv')
```

## 1
#### Geom_point makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types in Rhode Island by: 
```{r}
df %>% filter(Province_State=='Rhode Island') %>% ggplot()+geom_point(mapping = aes(x = Date, y = Doses_admin, color = Vaccine_Type))
```

#### Do the following:
- Use geom_point to compare the the Moderna and Pfizer vaccine administered in California
- Use geom_point to compare the totall vaccine administered in Rhode Island and Massachusetts
- Change geom_point to geom_line in the above codes to have the corresponding line plots.

```{r}
#first bullet point
df %>% filter(Province_State=='California') %>% ggplot()+geom_point(mapping = aes(x = Date, y = Doses_admin, color = Vaccine_Type))

#second bullet point
df %>% filter(Province_State==c('Rhode Island','Massachusetts')) %>% ggplot()+geom_point(mapping = aes(x = Date, y = Doses_admin, color = Vaccine_Type))

#the bottom two is for the third 
df %>% filter(Province_State==c('Rhode Island','Massachusetts')) %>% ggplot()+geom_line(mapping = aes(x = Date, y = Doses_admin, color = Vaccine_Type))

df %>% filter(Province_State=='California') %>% ggplot()+geom_line(mapping = aes(x = Date, y = Doses_admin, color = Vaccine_Type))
```


## 2
#### Sometime we need to do some calculations on the data before plotting. For example, to compare the popularity of two vaccine Moderna and Pfizer by date, we can use.

```{r}
# Create a dataset recording these two vaccine administered by date
d1 <- df %>% filter(Vaccine_Type==c('Moderna','Pfizer')) %>% group_by(Date, Vaccine_Type) %>% summarise(vac_admin=mean(Doses_admin, na.rm=TRUE))

# Plot from the newly create data
d1 %>% ggplot()+geom_point(mapping=aes(x=Date, y=vac_admin, color=Vaccine_Type))
```


#### Do the following: 
- Use geom_point to compare the popularity of all vaccines by date.
- Use geom_line to compare the popularity of all vaccines in New England by date.
- Use geom_line to compare the popularity of all vaccines in New England by month.
- Use geom_line to compare states in New England in term of the total vaccine administered by date.
- Use geom_line to compare states in New England in term of the total vaccine administered by date.
```{r}
#first bullet
d2 <- df %>% group_by(Date, Vaccine_Type) %>% summarise(vac_admin=mean(Doses_admin, na.rm=TRUE))

d2 %>% ggplot()+geom_point(mapping = aes(x = Date, y = vac_admin, color = Vaccine_Type))

#second bullet
d3 <- df %>% filter(Province_State==c('Rhode Island','Massachusetts','Vermont','Connecticut','New Hampshire', 'Maine')) %>% group_by(Date, Vaccine_Type) %>% summarise(vac_admin=mean(Doses_admin, na.rm=TRUE))

d3 %>% ggplot()+geom_line(mapping = aes(x = Date, y = vac_admin, color = Vaccine_Type))

#third bullet

#make a month column
library(lubridate)
df$Date <- ymd(df$Date)
df$month = month(df$Date)

d3 <- df %>% filter(Province_State==c('Rhode Island','Massachusetts','Vermont','Connecticut','New Hampshire', 'Maine'), Vaccine_Type==c('All')) %>% group_by(month, Vaccine_Type,Province_State) %>% summarise(vac_admin=mean(Doses_admin, na.rm=TRUE))

d3 %>% ggplot()+geom_line(mapping = aes(x = month, y = vac_admin,color = Province_State))

#forth bullet point

d4 <- df %>% filter(Province_State==c('Rhode Island','Massachusetts','Vermont','Connecticut','New Hampshire', 'Maine'),Vaccine_Type==c('All')) %>% group_by(Date, Vaccine_Type) %>% summarise(vac_admin=mean(Doses_admin, na.rm=TRUE))

d4 %>% ggplot()+geom_point(mapping = aes(x = Date, y = vac_admin, color = Vaccine_Type))

```













