---
title: "assignment 3"
author: "Nathan Angell"
date: "9/15/2021"
output: html_document
---
## problem 1

```{r}
df = data.frame(Rank = c(0,1,2,3),
                Age = c(28,34,29,42),
                Name = c("Tom", "Jack", "Steve", "Ricky"))
```
## problem 2
#### use the read.csv to get the df into your system
```{r}
df = read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
```

## problem 3
#### How many columns and rows the data have?
```{r}
str(df)
# there are 30,869 rows and 5 columns from what I saw from this code
```
## problem 4
#### How many missing values are there? Show the missing values by columns. What variable has the most number of missing values?
```{r}
colSums(is.na(df))
# there are no missing values
```
## problem 5
#### What is the class of the date column? Change the date columns to date type using the as.Date function. Show the new class of the date column
```{r}
class(df$date)

df$date <- as.Date(df$date)

```
## problem 6
####Capitalize the names of all the variables
```{r}
colnames(df) <- c('Date','State','Fips', 'Cases', 'Deaths')

```
## problem 7
#### Find the average number of cases per day. Find the maximum cases a day.
```{r}
# average number cases per day
mean(df$Cases, na.rm=TRUE)

#max cases a day
max(df$Cases, na.rm=TRUE)
```
## problem 8
#### How many states are there in the data?
```{r}
length(unique(df$State))
```
## problem 9
#### Create a new variable weekdays to store the weekday for each rows.
```{r}
df$Weekdays <- weekdays(df$Date)
```
## problem 10
####     Create the categorical variable death2 variable taking the values as follows: has_death if there is a death that day and no_death if there is no death that day
```{r}
df$Deaths2 <- ifelse(df$Deaths==0, 
                        'no_death','has_death ')
```
## problem 11
#### Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable death
```{r}
summary(df$Deaths)
```
## problem 12
#### Create the categorical variable death3 variable taking the values as follows: low_death if the number of deaths smaller than the 25 percentile (Q1) AND mid_death if the number of deaths from Q1 to Q3 AND high_death if the number of deaths greater than Q3
```{r}
df$Deaths3 <- cut(df$Deaths, 
                      breaks = c(0, 363, 7371,68106), 
                    labels=c('low_death','mid_death','high_death'))
table(df$Deaths3)
```
## problem 13
#### Find the average cases in Rhode Island in 2021
```{r}
# do not forget about the comma!
df1 <- df[(df$State == 'Rhode Island'),]

# then do the mean
mean(df1$Cases)
```
## problem 14
#### Find the median cases by weekdays in Rhode Island in 2021
```{r}
by(df1$Cases,df1$Weekdays, median)
```
## problem 15
#### Compare the median cases in Rhode Island in June, July, August and September in 2021.
```{r}
# it was this df2 <- df1[(df1$Date>='2021-06-01') & (df1$Date<='2021-09-01'),] but i changed it because we are including september

#seperate the dates above June 1st 2021
df2 <- df1[(df1$Date>='2021-06-01'),]

#create month column to seperate the values easily
df2$Month <- months(df2$Date)

# now use the by function to look at the median cases by each month
by(df2$Cases,df2$Month, median)

```
















