---
title: "Assignment 9"
author: "Nathan Angell"
date: "10/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## Import data and certain functions
```{r}
library(gganimate)
library(gifski)
library(tidyverse)
library(dplyr)
library(lubridate)

df <- read_csv("WHO-COVID-19-global-data.csv")
```

##  Find the three countries with the most numbers of deaths by Covid-19.
#### Hint: 1. Use filter to select the most current date. 2. Pipe with arrange(-Cumulative_deaths) to rank the data by the number of cumulative deaths, Make a plot revealing the number of deaths in the three countries with the most numbers of deaths. 3. Create the new variable (column) death_per_cases recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases). What are the three countries with the highest deaths per cases. 4. Make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 5. Use transition_reveal to create three animated graph with the vaccination data at

```{r}
# this is for #1 and #2

#I figured out my countries for this one
df1 <- df %>% filter(Date_reported >= "2021-10-17") %>% arrange(-Cumulative_deaths)

#create graph here
df %>% filter(Country %in% c('United States of America', 'Brazil', 'India')) %>% ggplot(aes(y=Cumulative_deaths,
            x=Date_reported,
            color=Country))+
  geom_line()+ 
    geom_point(size=3)+
    geom_text(aes(label = Cumulative_deaths), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)
```


```{r}
#this #3
df$death_per_cases <- df$Cumulative_deaths / df$Cumulative_cases

#look at the countries
df2 <- df %>% filter(Date_reported >= "2021-10-17") %>% arrange(-death_per_cases)

#answer for #3
# it is Yemen, Peru, and Mexico
```


```{r}
#this is for #4
df %>% filter(Country %in% c('United States of America', 'Italy', 'Mexico')) %>% ggplot(aes(y=death_per_cases,
            x=Date_reported,
            color=Country))+
  geom_line()+ 
    geom_point(size=3)+
    geom_text(aes(label = death_per_cases), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)
```


```{r}
#this is for #5
vac_df <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/vaccine_data_us_timeline.csv")

#first graph
vac_df %>% filter(Province_State %in% c('Rhode Island', 'Connecticut', 'Massachusetts')) %>% filter(Vaccine_Type == 'All') %>% group_by(Date) %>% ggplot(aes(y=Doses_shipped,
            x=Date,
            color=Province_State))+
  geom_line()+ 
    geom_point(size=3)+
    geom_text(aes(label = Doses_alloc), 
            hjust = -.1, size=5) +
  transition_reveal(Date)

#second graph
vac_df %>% filter(Province_State %in% c('California', 'Oregon', 'Washington')) %>% filter(Vaccine_Type == 'All') %>% group_by(Date) %>% ggplot(aes(y=Doses_shipped,
            x=Date,
            color=Province_State))+
  geom_line()+ 
    geom_point(size=3)+
    geom_text(aes(label = Doses_alloc), 
            hjust = -.1, size=5) +
  transition_reveal(Date)

#third graph
vac_df %>% filter(Province_State %in% c('Nebraska', 'Kansas', 'Oklahoma')) %>% filter(Vaccine_Type == 'All') %>% group_by(Date) %>% ggplot(aes(y=Doses_shipped,
            x=Date,
            color=Province_State))+
  geom_line()+ 
    geom_point(size=3)+
    geom_text(aes(label = Doses_alloc), 
            hjust = -.1, size=5) +
  transition_reveal(Date)

```


